---
title: "Haystack Cluster Analysis Notebook Part 2"
output: html_notebook
---
# Cluster Analysis - Haystack Data
The notebook uses clean, processes data from Step 3 - Data Augmentation Python
notebook
** Load Key Libraries **

```{r}
library(magrittr)
library(HDclassif)
library(psych)
library(cluster)
library(ggplot2)
library(tidyverse)
library(FactoMineR)
options(scipen=999)
```

### This is part 2 of Cluster Analysis, we will now use mixed data types to generate clusters


### Run an example of cluster analysis using FAMD
### Step1: FAMD
```{r}
#run_pca <- function(data_frame, pca_type, components)
dfl_famd <- run_pca(dfz_FAMD, "FAMD", 22)
```


```{r}
meth_list <- list("ward.D", "centroid","median")#, "average", "median", "centroid")
dist_list <- list("minkowski","maximum")#,"canberra")
bclist <- list()
# loop 
for (m in meth_list) {
  for (d in dist_list){
    best_cluster <- fun_nc(dfl_famd, d, 2, 5, m)
    #bclist  <- c(bclist, best_cluster)
    bclist[[length(bclist) + 1]] <- best_cluster
  }
}
```

Step 4: Execute the cluserting using dfl_FAMD
```{r}
#fun_clust <- function(data_frame, distance, agglo_method, clusters)
famd_min_cen_c <- fun_clust(dfl_famd, "minkowski", "ward.D", 4)
```

Evaluate Clusters
```{r}
table(famd_min_cen_c)
```

```{r}
tmp_data <- zdata
tmp_data$cluster <-famd_min_cen_c
tmp_data$cluster <- as.factor(tmp_data$cluster)
p <- ggplot(tmp_data, aes(x=cluster, y=rentZestimate)) + 
  geom_boxplot()
p
q <- ggplot(tmp_data, aes(x=cluster, y=Sch_Rat_Avg)) + 
  geom_boxplot()
q
p <- ggplot(tmp_data, aes(x=cluster, y=Income_per_return)) + 
  geom_boxplot()
p
q <- ggplot(tmp_data, aes(x=cluster, y=violent_crime_total_rate)) + 
  geom_boxplot()
q

```

### Step 5: Save the data to a file for exploratory analysis
```{r}
tmp_data$ClusterCategory <- ifelse(tmp_data$cluster == 1, "Cluster 1",
                              ifelse(tmp_data$cluster == 2, "Cluster 2",
                                     ifelse(tmp_data$cluster == 3, "Cluster 3",
                                            ifelse(tmp_data$cluster == 4, "Cluster 4",
                                     "Cluster 5"))))

write.csv(tmp_data,"cluster_output4.csv", row.names = FALSE)
```

Remove the temporary dataset
```{r}
rm(tmp_data)
```

### Extra evaluation using gower distance and PAM 

```{r}

gower_dist <- cluster::daisy(dfl_famd, metric = "gower")
```

```{r}
set.seed(123)
pam_cluster <- cluster::pam(gower_dist, k = 3)

table(pam_cluster$clustering)
```


```{r}
tmp_data <- zdata
tmp_data$cluster <-pam_cluster$clustering
tmp_data$cluster <- as.factor(tmp_data$cluster)
p <- ggplot(tmp_data, aes(x=cluster, y=livingArea)) + 
  geom_boxplot()
p
q <- ggplot(tmp_data, aes(x=cluster, y=Sch_Rat_Avg)) + 
  geom_boxplot()
q
p <- ggplot(tmp_data, aes(x=cluster, y=Income_per_return)) + 
  geom_boxplot()
p
q <- ggplot(tmp_data, aes(x=cluster, y=violent_crime_total_rate)) + 
  geom_boxplot()
q

```

*** End of Part 2 ***